## Chernoff’s inequality

The general Hoeffding's inequality are not so sharp in case $(X_{i})_{i \in \mathbb{N}^*}$ are Bernoulli with parameter $p_i$ small enough to apply Poisson Limit Theorem: suppose $\mathop{{}\mathbb{E}}_{}S_n=\sum_{}^{} p_i=\mu$, then by Hoeffding's inequality \@ref(thm:general-hoeffding):
$$
\mathop{{}\mathbb{P}}\left\{ S_n-\mu \ge t \right\}\le \exp(- \frac{2t^2}{n})
$$
which is not so sharp as Chernoff’s inequality giving below, since it's insensitive to the magnitude of $p_i$ (and thus, $\mu$). Figure \@ref(fig:hoeffding-chernoff) shows their difference.

```{r hoeffding-chernoff, fig.align = "center", fig.cap="behavior when $\\mu=1$", out.width = '70%'}

knitr::include_graphics("figures/hoeffding-chernoff.pdf")

```

::: {.theorem #chernoff-inequality name="chernoff inequality"}

Let $X_{i}$ be independent Bernoulli r.v. with parameter $p_{i}$. Consider their sum $S_n=\sum_{i=1}^{n} X_{i}$ and denote $\mu=\mathbb{E}S_n$ then for $t\gt \mu$ we have:
$$ 
\mathbb{P}\left\{S_n\ge t\right\}\le e^{t-\mu}\left(\frac{\mu}{t}\right)^{t}
$$

:::

::: {.proof}

By Chernoff bound and independency of $X_i$:
$$
\mathop{{}\mathbb{P}}\left\{ S_n \ge t \right\}\le e^{-\lambda t} M_{S_n}(t)=e^{-\lambda t} \prod_{i=1}^{n} M_{X_i}(\lambda)
$$
Note the numeric inequality $1+x \le e^{x}$:
$$ 
\mathbb{E}\exp(\lambda X_{i})=e^{\lambda}p_i+(1-p_i)=1+\left(e^{\lambda}-1\right)p_i\le \exp\left((e^{\lambda}-1)p_i\right) 
$$
Then
$$
\prod_{i=1}^{N} \mathbb{E} \exp(\lambda X_{i})\le \exp\left[(e^{\lambda}-1)\sum_{i=1}^{N} p_i\right]=\exp[(e^{\lambda}-1)\mu]
$$
Finally, pick $\lambda=\ln(t /\mu)$, we complete the proof.

:::

::: {.proposition  name="lower tail"}

For any $t\lt \mu$ we have
$$
\mathbb{P}\left\{S_N\le t\right\}\le e^{t-\mu}\left(\frac{\mu}{t}\right)^{t}
$$

:::

::: {.proof}

Note:
$$
\begin{aligned}
    \mathbb{P}\left\{S_N\le t\right\}&= \mathbb{P}\left\{-S_N\ge -t\right\} \le \frac{\mathbb{E}\exp(-\lambda S_N)}{\exp(-\lambda t)}  \\
    &= e^{\lambda t}\cdot \prod_{i=1}^{N} \mathbb{E}e^{-\lambda X_{i}} \le e^{\lambda t} \exp \left[ (e^{-\lambda}-1)\mu \right]
\end{aligned}
$$
Then replace $-\lambda=\ln(t /\mu)$.

:::

::: {.proposition  name="Poisson tail"}

Let $X\sim \mathop{\text{Pois}}(\lambda)$, then for any $t\gt \mu$, we have
$$
\mathbb{P}\left\{X\ge t\right\}\le e^{-\lambda}\left(\frac{e \lambda}{t}\right)^{t}
$$

:::

::: {.proof}

According to the Poisson limit theorem, if $X\sim \mathop{\text{Pois}}(\lambda)$, we can construct a series of r.v. where $X_{i}\sim \mathop{\text{Ber}}(p_{i})$ s.t. $X=\sum_{i=1}^{\infty} X_{i}$ and note that $X_{i}\ge 0$ for every $i$ so we can use MCT.
$$
\begin{aligned}
    \mathbb{P}\left\{X\ge t\right\}&\le 
    \exp(-\lambda t)\cdot \prod_{i=1}^{\infty} \mathbb{E}\exp(\lambda X_{i})
    \\ & \le 
    \exp \left( \lambda(e^{t}-1) \right)
\end{aligned}
$$
It follows some similar step and we done.

:::


::: {.proposition #chernoff-inequality-small name="Chernoff’s inequality: small deviations"}

In the setting of theorem \@ref(thm:chernoff-inequality), for $\delta \in (0,1]$, we have
$$
\mathop{{}\mathbb{P}}\left\{ \left| S_n-\mu \right|\ge \delta\mu \right\} \le 2 \exp \left( -c \mu \delta^2 \right)
$$
where $c$ is an absolute constant.

:::


::: {.proof}

We only show the positive side, by chernoff inequality
$$
\begin{aligned}
    \mathop{{}\mathbb{P}}\left\{ S_n-\mu \ge \delta\mu \right\}&\le e^{-\lambda (\delta+1)\mu} \mathop{{}\mathbb{E}}_{}\exp \left( \lambda S_n \right)
    \le e^{\delta \mu} \left( \frac{1}{\delta+1} \right)^{(\delta+1)\mu}
\end{aligned}
$$
Taking log each side, then we show that
$$
-\delta+(\delta+1) \ln(\delta+1) \ge  c\delta^2
$$
holds for some $c$. It's possible by noting 

```{r chernoff-small, fig.align = "center", fig.cap=" small deviations", out.width = '70%'}

knitr::include_graphics("figures/chernoff-small.pdf")

```

:::

As consequence, by Poisson Limit Theorem, we have when $t$ is small, Poisson is almost sub-gaussian(see later):

::: {.corollary  name=""}

Let $X\sim \text{Pois}(\lambda)$, for $t \in (0,\lambda]$, we have
$$
\mathop{{}\mathbb{P}}\left\{ \left| X-\mu \right|\ge t \right\}\le 2\exp \left( - \frac{ct^2}{\lambda} \right)
$$

:::



<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
